---
title: "Activity Monitoring Data"
author: "Guilherme Ramos Casimiro"
date: "February 19, 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(out.extra = 61)
```

## Loading the required libraries

```{r library, message = FALSE}
library(dplyr)
library(lattice)
```


## Download the dataset


```{r cars}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, "data.zip")
unzip("data.zip")
```

## Read and clean the data



```{r pressure}
activity <- read.csv("activity.csv")
activity <- na.omit(activity)
activity$date <- as.Date(activity$date)
group.activity <- group_by(activity, date) %>% summarise(total = sum(steps),
                                                         mean = mean(steps), median = median(steps))
```

## Plot the histogram of total steps per day

```{r hist}
histogram(group.activity$total, main = "Total number of steps per day", xlab = "Number of total steps")
```

## Show the mean of steps taken each day 

The transpose shows all the data groups.

```{r mean}
t(group.activity[,3:4])
```

## Plot time series: mean of steps per interval across all the days

```{r time series}
group.5minute <- group_by(activity, interval) %>% summarise(mean = mean(steps, na.rm = TRUE))
with(group.5minute, plot(interval, mean, type = "l",
                         main = "Mean of steps per interval across all the days",
                         xlab = "Interval", ylab = "Mean of steps"))
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r mean interval}
filter(group.5minute,mean == max(group.5minute$mean))[1]
```

## Imputing missing values

The number of missing values in this dataset is 2304

```{r missing sum}
activity.na <- read.csv("activity.csv")
sum(is.na(activity.na))
```

## Imputing missing data solution

The strategy for filling in all of the missing data is to fill of zeros, because all group of missing data corresponds to a specific group of dates, e.g, the date 2012-10-01 contains only missing data to steps field.

```{r missing}
activity.na$date <- as.Date(activity.na$date)
na_steps <- is.na(activity.na$steps)

activity.na[na_steps,1] <- 0
group.activity.na <- group_by(activity.na, date) %>% summarise(total = sum(steps),
                                                         mean = mean(steps), median = median(steps))
```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r hist missing}

histogram(group.activity.na$total, main = "Total number of steps per day after missing values imputed", xlab = "Number of total steps")
```

The transpose shows all the data groups.

```{r mean missing}
t(group.activity.na[,3:4])
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r weekdays}
weekdays <- weekdays(activity.na$date, TRUE)
f2 <- function(x) if(x %in% c("Sat","Sun")) "Weekend" else "Weekday"
activity.na$week.factor <- sapply(weekdays, f2)
group.activity.na <- group_by(activity.na, interval, week.factor) %>% summarise(mean = mean(steps))
xyplot(mean~interval | week.factor,
       data = group.activity.na,
       layout = c(1,2),
       xlab = "Interval",
       ylab = "Number of steps",
       type = "l")
```